---
title: "Untitled"
author: "Terry Lin"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
fna.data <- read.csv("C:/Users/Terry/Desktop/R/class09/WisconsinCancer.csv")

wisc.df <- fna.data
x<- table(wisc.df$diagnosis)

#length of grep vector, value=true gives the values of those that are true
length(grep("\\_mean",colnames(wisc.df), value=TRUE))
```
Examine data from `r nrow(wisc.df)` patient samples.
In this dataset, we have `r x["M"]` cancer and `r x["B"]` non-cancer.

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
row.names(wisc.data) <- wisc.df$id
diagnosis <- wisc.df$diagnosis
```

```{r}

round(apply(wisc.data,2,sd),3)
```

We will need to set scale to be true if standard deviation is very large

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```
Can look at individual PC % or cumulative to cover 90% of original variance in data

```{r}
biplot(wisc.pr)
```


Plot PCA with PC1 AND PC2 by diagnosis
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis, xlab= "PC1", ylab="PC2")
```


```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,3], col=diagnosis, xlab="PC1", ylab="PC3")
```

Calculating variance of each PC
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)

#need to divide each variance in the vector by the sum of the vector
pve <- pr.var/sum(pr.var)

plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", ylim= c(0,1), type= "o")
```


```{r}
x <- summary(wisc.pr)
x$importance[,"PC1"]
```

PC1 captures `r x$importance[2,"PC1"] *100` of the original variance in the dataset.



Alternate scree plot of the same data
```{r}
barplot(pve, ylab="Percent of Variance Explained", names.arg=paste0("PC", 1:length(pve)), las=2, axes=FALSE)
axis(2, at=pve, labels=round(pve,2)*100)
```
ggplot based graph
```{r}
library(factoextra)
fviz_eig(wisc.pr, addlabels=TRUE)
```


```{r}
wisc.pr$rotation["smoothness_se",1]
```



Hierarchical Clustering

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist)

plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```


other way to do hierarchical clustering, using the outcome of the PCA 

```{r}
other<-hclust(dist(wisc.pr$x[,1:3]))
plot(other)

```


```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters, diagnosis)
```

___________________________________________________________________________________

K-means Clustering

```{r}
wisc.km <- kmeans(wisc.data, centers=2, nstart=20)
table(wisc.km$cluster, diagnosis)
```


Combining Methods

```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), "ward.D2")
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps,diagnosis)
plot(wisc.pr$x[,1:2], col=grps)
plot(wisc.pr$x[,1:2], col=diagnosis)
```

Turn our groups into a factor and reorder the levels.
```{r}
g <- as.factor(grps)
levels(g)
g <- relevel (g,2)
levels(g)

plot(wisc.pr$x[,1:2], col=g)
```

3D package rgl

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC1", ylab="PC2", zlab="PC3", cex=1.5, size=1, type="s", col=grps)
rglwidget(width=400,height=400)
```
rgl


```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
table(wisc.pr.hclust.clusters,diagnosis)
table(wisc.km$cluster, diagnosis)

```




Making predictions

```{r}
new <- read.csv("C:/Users/Terry/Desktop/R/class09/new_samples.csv")
npc <- predict(wisc.pr, newdata=new)
npc

plot(wisc.pr$x[,1:2],col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```






